Analysis

### 2023-02-17 Analysis.rmd
test commit to test overleaf sync

colors <- RColorBrewer::brewer.pal(n = 9, name = "Set1")

@@

models_all <- forecasts$model |> unique() |> sort()
models_all <- models_all[!models_all %in% c("", "EpiExpert-ensemble")]
models_non_human <- c("crowd-ensemble", "crowd-direct", "crowd-rt",
                      "EuroCOVIDhub-ensemble")
models_human <- models_all[!(models_all %in% models_non_human)]
models_rt <- models_human[grepl("(Rt)", models_human)]
models_direct <- models_human[!grepl("(Rt)", models_human)]

@@

```{r}
plot_data <- function(weekly_truth, daily_truth) {
  weekly_truth |>
  ggplot(aes(x = target_end_date, y = true_value)) +
  annotate("rect", xmin = time_start, xmax = time_stop, ymin = 0, ymax = Inf,
           fill = "darkolivegreen3", alpha = .1) + 
  geom_line() + 
  geom_point(size = 0.8, aes(color = status)) + 
  geom_bar(data = daily_truth |>
             mutate(true_value = true_value * 7), 
           stat = "identity", 
           fill = "skyblue3", alpha = 0.4) + 
  facet_wrap(~ target_type, 
             nrow = 2, scale = "free_y") + 
  scale_y_continuous(labels = label_fn) +
  theme_scoringutils() + 
  scale_colour_manual(values = c("red", "black")) + 
  labs(y = "Observed value", x = "Date", color = "Data status")
}

plot_data_revised <- plot_data(weekly_truth_revised, daily_truth_revised)
plot_data <- plot_data(weekly_truth, daily_truth)

ggsave("output/figures/plot-data-revised.png", plot = plot_data_revised, 
       width = 7, height = 5)

ggsave("output/figures/plot-data.png", plot = plot_data, 
       width = 7, height = 3.5)

Show data revisions

plot_weekly_revisions <- weekly_truth |>
  rbind(weekly_truth_revised) |>
  filter(target_type == "Deaths") |>
  select(target_end_date, true_value, target_type) |>
  group_by(target_end_date, target_type) |>
  mutate(diff = c(NA, diff(true_value))) |>
  filter(!is.na(diff)) |>
  ggplot(aes(x = target_end_date, y = diff)) + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80") + 
  geom_line() + 
  geom_point(size = 0.7) + 
  facet_wrap(~ target_type, ncol = 1, scale = "free_y") +
  labs(y = "Difference in data", x = "Date") + 
  coord_cartesian(ylim = c(-500, 500)) + 
  theme_scoringutils()

ggsave("output/figures/data-revisions.png", plot = plot_weekly_revisions, 
       width = 7, height = 2.5)

Participant forecasts 
```{r}
models <- forecasts$model |> unique()

plot_df <- forecasts |>
  filter(quantile == 0.5) |>
  rbind(horizon_0) |>
  filter(target_end_date >= "2021-04-01", 
         target_end_date <= "2021-10-01") |>
  filter(!(model %in% c("crowd-rt", "EpiExpert-ensemble"))) |>
  filter(!(grepl("(Rt)", model))) |>
  mutate(colouring = ifelse(model == "crowd-ensemble", 
                            "Crowd ensemble", 
                            "individual particpant"))

Median forecasts ensemble
```{r}

forecasts |>
  filter(quantile == 0.5) |>
  rbind(horizon_0) |>
  filter(target_end_date >= "2021-04-01", 
         target_end_date <= "2021-10-01") |>
  filter((model %in% c("crowd-ensemble", "EuroCOVIDhub-ensemble"))) |>
  ggplot(aes(x = target_end_date)) +
  geom_line(aes(y = prediction, group = interaction(forecast_date, model), 
                color = model), 
            linewidth = .4) + 
  geom_line(aes(y = true_value),
            linewidth = .4, color = "black") + 
  geom_point(aes(y = true_value),
             size = .5, color = "black") + 
  scale_color_manual(values = colors[1:2]) +
  facet_wrap(~ target_type, scale = "free_y", ncol = 1) + 
  labs(x = "Date", y = "Median forecast and observed values", color = "Forecaster") + 
  theme_scoringutils()
```

comparison_plot <- function(scores, models = c("crowd-direct", "EuroCOVIDhub-ensemble")) {
  mean_wis <- scores |>
  filter(scale %in% c("log", "natural")) |>
  filter(model %in% models) |>
  summarise_scores(by = c("model", "scale", "target_type", "type_and_scale")) |>
  summarise_scores(by = c("model", "scale", "target_type", "type_and_scale"),
                   fun = signif, digits = 3) |>
  mutate(wis_mean = paste("Mean:", label_fn(interval_score)))

## Weekly participation

@@

avail_ensemble <- ensemble_inclusions |>
  mutate(prediction = 1, true_value = 1) |>
  avail_forecasts(by = c("forecast_date", "target_type"), 
                    collapse = c("horizon", "quantile", "target_end_date")) |>
  mutate(group = "Forecast Hub")

@@

  cov_plot <- function(scores, filter_range) {
    scores |>
      filter(range == filter_range) |>
      summarise_scores(by = c("model", "target_type", "range", "horizon")) |>
      ggplot(aes(x = horizon, y = coverage, color = model)) + 
      geom_line(show.legend = FALSE) + 
      geom_point(show.legend = FALSE) + 
      geom_hline(yintercept = filter_range / 100, 
                 linetype = "dashed", colour = "grey80") + 
      facet_wrap(~ target_type, ncol = 2) + 
      scale_y_continuous(labels = \(x) {paste0(100 * x, "%")}) + 
      theme_scoringutils() + 
      labs(x = "Horizon", y = paste("Coverage -", filter_range, "PI"))
  }

@@

ggsave("output/figures/performance.png", plot = p_performance, 
       width = 7, height = 8)

@@

ggsave("output/figures/comparison-direct-rt-individual.png", 
       plot = p_comparison_direct_rt_individual, 
       width = 7, height = 5)

@@

```{r, performance-table}
make_performance_table <- function(scores) {
  scores |>
    filter(scale %in% c("log", "natural")) |>
    summarise_scores(by = c("model", "target_type", "scale", "horizon")) |>
    filter(horizon == 2) |>
    select(model, target_type, scale, interval_score, coverage_50, coverage_90) |>
    pivot_wider(names_from = scale, values_from = interval_score) |>
    select(model, target_type, natural, log, coverage_50, coverage_90) |>
    mutate(across(c(where(is.numeric)), signif, digits = 2)) |>
    kable(format = "latex", booktabs = TRUE, 
          align = c("l", "l", rep("c", 4)),
          caption = "Performance for two-week-ahead forecasts \\label{tab:scores}",
          linesep = c('', '', '', '\\addlinespace'),
          col.names = c("Model", "Target", "WIS", "WIS (log scale)", 
                        "Coverage 50%", "Coverage 90%")) |> 
    kable_styling(latex_options = c("scale_down", 
                                    "hold_position")) 
}

@@

@@

Methods
- Hub ensemble switched from mean to median on 2021-07-19
- did we do do mean or median? 
- how did we end up with the crowd ensemble

>>>>

colors <- RColorBrewer::brewer.pal(n = 9, name = "Set1")
```

# Define Helper Functions, Load Data

@@

models_all <- forecasts$model |> unique() |> sort()
models_all <- models_all[!models_all %in% c("", "EpiExpert-ensemble")]
models_non_human <- c("crowd-ensemble", "crowd-direct", "crowd-rt",
                      "EuroCOVIDhub-ensemble")
models_human <- models_all[!(models_all %in% models_non_human)]
models_rt <- models_human[grepl("(Rt)", models_human)]
models_direct <- models_human[!grepl("(Rt)", models_human)]
models_expert <- forecasts |>
  filter(expert) |>
  pull(model) |> unique() |> sort()

update_expert_status <- function(data) {
  data |>
    mutate(expert = ifelse(
      model %in% models_expert | model %in% paste(models_expert, "(Rt)"), 
      TRUE, 
      FALSE
    ))
}

@@

# Figure SI.1: Data and data revisions
data revisions: https://github.com/CSSEGISandData/COVID-19/issues/6021
https://coronavirus.data.gov.uk/details/whats-new 
https://coronavirus.data.gov.uk/details/whats-new/record/965a7923-dacf-42ed-abb9-3a972dbf1835

```{r}
plot_data <- function(weekly_truth, daily_truth) {
  weekly_truth |>
    mutate(status = str_to_title(status)) |>
  ggplot(aes(x = target_end_date, y = true_value)) +
  annotate("rect", xmin = time_start, xmax = time_stop, ymin = -Inf, ymax = Inf,
           fill = "darkolivegreen3", alpha = .1) + 
  geom_line() + 
  geom_point(size = 0.8, aes(color = status)) + 
  geom_bar(data = daily_truth |>
             mutate(true_value = true_value * 7), 
           stat = "identity", 
           fill = "skyblue3", alpha = 0.4) + 
  facet_wrap(~ target_type, 
             nrow = 2, scale = "free_y") + 
  scale_y_continuous(labels = label_fn) +
  theme_scoringutils() + 
  scale_colour_manual(values = c("red", "black")) + 
  labs(y = "Original data", x = "Date", color = "Data status")
}

p_data_revised <- plot_data(weekly_truth_revised |>
                                 filter(target_type == "Deaths"), 
                               daily_truth_revised |>
                                 filter(target_type == "Deaths")) + 
  ylab("Revised data")
p_data <- plot_data(weekly_truth, daily_truth)

p_weekly_revisions <- weekly_truth |>
  rbind(weekly_truth_revised) |>
  filter(target_type == "Deaths") |>
  select(target_end_date, true_value, target_type) |>
  group_by(target_end_date, target_type) |>
  mutate(diff = c(NA, diff(true_value))) |>
  filter(!is.na(diff)) |>
  ggplot(aes(x = target_end_date, y = diff)) + 
  annotate("rect", xmin = time_start, xmax = time_stop, ymin = -Inf, ymax = Inf,
           fill = "darkolivegreen3", alpha = .1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80") + 
  geom_line() + 
  geom_point(size = 0.7) + 
  facet_wrap(~ target_type, ncol = 1, scale = "free_y") +
  labs(y = "Weekly data revisions", x = "Date") + 
  coord_cartesian(ylim = c(-500, 500)) + 
  theme_scoringutils()

p_all_data <- p_data / 
  p_data_revised / 
  p_weekly_revisions +
  plot_annotation(tag_levels = "A") + 
  plot_layout(guides = "collect", heights = c(2.2, 1, 1)) & 
  theme(legend.position = "bottom")

p_all_data

ggsave("output/figures/plot-data.png", plot = p_all_data, 
       width = 7, height = 8)
```

# Figure SI.4: Weekly participation humans and hub ensemble

@@

avail_ensemble <- ensemble_inclusions |>
  mutate(prediction = 1, true_value = 1) |>
  filter(forecast_date <= "2021-08-16") |>
  unique() |>
  avail_forecasts(by = c("forecast_date", "target_type"), 
                    collapse = c("horizon", "quantile", "target_end_date")) |>
  mutate(group = "Forecast Hub")

@@

  cov_plot <- function(scores, filter_range) {
    scores |>
      filter(range == filter_range) |>
      summarise_scores(by = c("model", "target_type", "range", "horizon")) |>
      ggplot(aes(x = horizon, y = coverage, color = model)) + 
      geom_line(show.legend = FALSE) + 
      geom_point(show.legend = FALSE, size = 0.7) + 
      geom_hline(yintercept = filter_range / 100, 
                 linetype = "dashed", colour = "grey80") + 
      facet_wrap(~ target_type, ncol = 2) + 
      scale_y_continuous(labels = \(x) {paste0(100 * x, "%")}) + 
      theme_scoringutils() + 
      labs(x = "Horizon", y = paste("Coverage -", filter_range, "PI"))
  }

@@

ggsave("output/figures/performance.png", plot = p_performance, 
       width = 7, height = 6)

@@

ggsave("output/figures/comparison-direct-rt-individual.png", 
       plot = p_comparison_direct_rt_individual, 
       width = 7, height = 4)

@@

```{r, performance-table}
make_performance_table <- function(scores) {
  scores |>
    filter(scale %in% c("log", "natural")) |>
    summarise_scores(by = c("model", "target_type", "scale", "horizon")) |>
    filter(horizon == 2) |>
    select(model, target_type, scale, interval_score, coverage_50, coverage_90) |>
    pivot_wider(names_from = scale, values_from = interval_score) |>
    select(model, target_type, natural, log, coverage_50, coverage_90) |>
    mutate(across(c(where(is.numeric)), signif, digits = 2)) |>
    arrange(target_type, model) |>
    kable(format = "latex", booktabs = TRUE, 
          align = c("l", "l", rep("c", 4)),
          caption = "Performance for two-week-ahead forecasts \\label{tab:scores}",
          linesep = c('', '', '', '\\addlinespace'),
          col.names = c("Model", "Target", "WIS", "WIS (log scale)", 
                        "Coverage 50%", "Coverage 90%")) |> 
    kable_styling(latex_options = c("scale_down", 
                                    "hold_position")) 
}

@@

## Look at expert forecasts

# ensemble of experts and non-experts

create_expert_ensemble <- function(forecasts) {

  forecasts_expert_ensemble <- forecasts |>
    filter(model %in% models_human) |>
    update_expert_status() |>
    group_by(target_type, horizon, expert, quantile, true_value, forecast_date, target_end_date) |>
    summarise(prediction = median(prediction), 
              n = n()) |>
    mutate(model = ifelse(expert, "Expert", "Non-Expert")) |>
    as.data.table()

  scores_expert_ensemble <- forecasts_expert_ensemble |>
    mutate(method = model) |>
    score_and_add_log()

  return(list(forecasts = forecasts_expert_ensemble, 
              scores = scores_expert_ensemble))
}

expert_ensemble <- create_expert_ensemble(filter(forecasts))

# plot with number of experts and non-experts every week
expert_ensemble$forecasts |>
  filter(horizon == 2) |>
  group_by(model, forecast_date, target_type) |>
  summarise(n = unique(n)) |>
  ggplot(aes(x = forecast_date, y = n, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_scoringutils() + 
  facet_wrap(~ target_type)

expert_ensemble$scores |>
   make_performance_table()

p_scores_expert <- plot_performance(expert_ensemble$scores)
ggsave("output/figures/performance-expert.png", 
       plot = p_scores_expert, 
       width = 7, height = 6)

expert_ensemble_direct <- forecasts |>
  filter(model %in% models_direct) |>
  create_expert_ensemble()

expert_ensemble_direct$scores |>
   make_performance_table()

p_scores_expert_direct <- plot_performance(expert_ensemble_direct$scores)
ggsave("output/figures/performance-expert-direct.png", 
       plot = p_scores_expert_direct, 
       width = 7, height = 6)

# plot_scores_and_pred(
#   forecasts_expert_ensemble,
#   scores_expert_ensemble
# )

```{r}
scores_df <- scores  |>
  update_expert_status() |>
  mutate(expert = ifelse(expert, "Expert", "Non-Expert")) |>
  filter(model %in% models_human) |>
  filter(horizon == 2) |>
  summarise_scores()

mean_wis <- scores_df |>
  group_by(scale, expert, target_type, type_and_scale) |>
  summarise(wis_mean = signif(median(interval_score), 3)) |>
  mutate(wis_mean = paste("Mean:", label_fn(wis_mean)))

scores_df |>   
  ggplot(aes(x = expert, y = interval_score)) +
  geom_violin(alpha = 0.2) + 
  geom_boxplot(alpha = 0.5) + 
  # scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  geom_text(data = mean_wis, 
            aes(label = unique(wis_mean), y = -Inf), 
            vjust   = -1) + 
  facet_wrap(~ type_and_scale, scale = "free_y") + 
  scale_y_continuous(labels = label_fn, 
                     expand = expansion(mult = c(0.15, 0.05))) + 
  theme_scoringutils()

scores_df |>
  summarise_scores(by = c("expert", "target_type", "scale", "horizon")) |>
  rename(model = expert) |>
  make_performance_table()

@@

Methods
- Hub ensemble switched from mean to median on 2021-07-19

# Parking lot
Median forecasts for a 1:4 horizon
```{r}
models <- forecasts$model |> unique()

plot_df <- forecasts |>
  filter(quantile == 0.5) |>
  rbind(horizon_0, fill = TRUE) |>
  filter(target_end_date >= "2021-04-01", 
         target_end_date <= "2021-10-01") |>
  filter(!(model %in% c("crowd-rt", "EpiExpert-ensemble"))) |>
  filter(!(grepl("(Rt)", model))) |>
  mutate(colouring = ifelse(model == "crowd-ensemble", 
                            "Crowd ensemble", 
                            "individual particpant"))

Median forecasts combined human ensemble and Forecast Hub ensemble 1:4 horizon
```{r}
forecasts |>
  filter(quantile == 0.5) |>
  rbind(horizon_0, fill = TRUE) |>
  filter(target_end_date >= "2021-04-01", 
         target_end_date <= "2021-10-01") |>
  filter((model %in% c("crowd-ensemble", "EuroCOVIDhub-ensemble"))) |>
  ggplot(aes(x = target_end_date)) +
  geom_line(aes(y = prediction, group = interaction(forecast_date, model), 
                color = model), 
            linewidth = .4) + 
  geom_line(aes(y = true_value),
            linewidth = .4, color = "black") + 
  geom_point(aes(y = true_value),
             size = .5, color = "black") + 
  scale_color_manual(values = colors[1:2]) +
  facet_wrap(~ target_type, scale = "free_y", ncol = 1) + 
  labs(x = "Date", y = "Median forecast and observed values", color = "Forecaster") + 
  theme_scoringutils()
```

## Violin plot to compare scores

```{r}
comparison_plot <- function(scores, models = c("crowd-direct", "EuroCOVIDhub-ensemble")) {
  mean_wis <- scores |>
  filter(scale %in% c("log", "natural")) |>
  filter(model %in% models) |>
  summarise_scores(by = c("model", "scale", "target_type", "type_and_scale")) |>
  summarise_scores(by = c("model", "scale", "target_type", "type_and_scale"),
                   fun = signif, digits = 3) |>
  mutate(wis_mean = paste("Mean:", label_fn(interval_score)))
